Sampling requirements to detect ecosystem change

### 2020-06-22 R/manuscript_draft.Rmd
changes to cover page

Number of words: XXXX \vspace{5 mm}

Number of references: <!--\total{citenum}-->

Number of figures: 5

@@

Submission as: Article at \emph{Journal of Applied Ecology}

@@

Preprint available at:

@@

\begin{center}
\textbf{\Large Sampling requirements and approaches to detect ecosystem shifts}
\vspace{5 mm}

\textsc{Rosalie Bruel$^{*1}$\footnote{*The authors contributed equally to this work. $\dagger$Corresponding author (eastonrwhite@gmail.com)} and Easton R. White$^{* \dagger 2,3}$}
\vspace{3 mm}

@@

<!--\linenumbers-->
\singlespacing

@@

\vspace{3 mm}

\textbf{Keywords:} time series, changepoint, monitoring, optimal sampling, Cladocera

@@

@@

However, recent work suggest that combining indicators (in the specific study, trait dynamics and abundance-based early warning signals) allows forecasating population collapses even with at lower resolution and time-series length [@Arkilanian2020]. Critical slowing down does not necessarily result in a shift, and a shift can occur without critical slowing down [@Spears2017]. Signs of critical slowing downs are important to understand and recognize because they provide potential early warnings [@Doncaster2016], but in terms of management, knowing the timing of a shift can have larger implications in addressing the underlying driver. Thus, selecting a set number of samples or specific approach may also limit what future questions can be asked.

>>>>

Number of words: 3400 \vspace{5 mm}

Number of references: 35<!--\total{citenum}-->

Number of figures: 6

@@

Submission as: Article at \emph{Ecological Applications}

@@

Preprint available at: https://www.biorxiv.org/content/10.1101/2020.04.07.030643v1

@@

\begin{center}
\textbf{\Large Sampling requirements and approaches to detect ecosystem shifts}
\vspace{2 mm}

\textsc{Rosalie Bruel$^{*1}$\footnote{*The authors contributed equally to this work. $\dagger$Corresponding author (eastonrwhite@gmail.com)} and Easton R. White$^{* \dagger 2,3}$}
\vspace{1 mm}

@@

\linenumbers
\doublespacing

@@

\vspace{0.5 mm}

\textbf{Keywords:} time series, changepoint, monitoring, optimal sampling, Cladocera, paleoecology

@@

However, recent work suggest that combining indicators (in the specific study, trait dynamics and abundance-based early warning signals) allows forecasting population collapses even with at lower resolution and time-series length [@Arkilanian2020]. Critical slowing down does not necessarily result in a shift, and a shift can occur without critical slowing down [@Spears2017]. Signs of critical slowing downs are important to understand and recognize because they provide potential early warnings [@Doncaster2016], but in terms of management, knowing the timing of a shift can have larger implications in addressing the underlying driver. Thus, selecting a set number of samples or specific approach may also limit what future questions can be asked.

# Code availability

Data and code for all the figures can be found at https://github.com/rosalieb/temporal-sampling.

<div id="refs"></div>

<!--# Figures-->

@@
